Untitled

### 2021-02-24 outputs/paper/paper.Rmd
Commit for discussion

---
title: "COVID Shutdown Regulations on Restaurants Decreases xx% of Average Total Sales in Ontario"
author: "Hong Shi, Hong Pan, Yixin Guan, Babak Mokri"
thanks: "Code and data are available at: https://github.com/honn-ishinn/restaurants_covid_shutdown"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Since 2020, the world-wide COVID-19 pandemic has greatly influenced public health and healthcare, resulting in governments to impose regulations on social gatherings to eliminate further outbreaks. In Ontario, the provincewide stay at home order and shutdown restrictions on businesses and facilities has limited economic activities into a recession. This paper examines the shutdown effect on restaurants by reopening a city in Ontario (Peterborough) and comparing its restaurant operating situation with that of another city (Brantford) after three months reopening. The experiment results show that shutdown regulations on restaurants have decresed xxx%....."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

@@

@@

This experiment considered all Ontario restaurants as its target population. The frame population was decided by randomly selecting 100 restaurants from Brantford and 100 from Peterborough. To ensure the external validity of the experiment, we used the following steps to find these restaurants:

@@

@@

* Since the distance between Brantford and Peterborough is more than 200 kilometers apart, residents in Brantford are unlikely to travel such long distance solely for a dine-in meal. So the treatment of reopening Peterborough would not affect the mealing behavior in Brantford. However, residents in shutdown regions near Peterborough might go for dine-in meals, increasing sales in Peterborough restaurants. The actual shutdown effect on restaurants could be overestimated due to possible increase in sales from nearby shutdown regions' residents.

* The mindful comparison between cities' similarities on employment, population density, etc. ensures the internal validity of our experiment so that we could accurately estimate the true impact of the shutdown restriction on restaurant. However, the shutdown effect evaluation may not be generalized to the entire population of interest, i.e. all restaurants in Ontario such as township regions or cosmopolitan city Toronto.

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Sampling Method

@@

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

>>>>

---
title: "COVID Shutdown Regulations Negatively Impact Employment and Sales of Restaurants in Ontario"
author: "Hong Shi, Hong Pan, Yixin Guan, Babak Mokri"
thanks: "Code and data are available at: https://github.com/honn-ishinn/restaurants_covid_shutdown"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Due to the world-wide COVID-19 pandemic, shutdown restrictions on businesses and facilities have limited economic activities into a recession. This paper examines the shutdown effect on restaurants by reopening a city in Ontario (Peterborough) and comparing its restaurant operating situation with that of another city (Brantford) after three months reopening. The survey result indicates that shutdown restrictions have negative impacts on restaurant employment, sales and future operation. However, the higher increased COVID cases in the reopened city sugguests the Ontario government to mindfully plan for futher policy decisions."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

@@

&nbsp; &nbsp; The COVID-19 pandemic has impacted many businesses, including restaurants, and resulted in mass-layoffs of workers throughout the world. According to HuffPost [@cite_unemployment], “Canada's unemployment rate was 13.5% in May 2020, the highest it has been since 1976” [@cite_wikiemploy]. However, the unemployment crisis in Canada, which would result in rising rates of depression, suicide, frustration, crime, and other unprecedented social issues, was proactively considered and managed by the Government of Canada. The Canada Emergency Wage Subsidy (CEWS) [@cite_CEWS] as well as the Canada Emergency Rent Subsidy (CERS) [@cite_CERS] programs were enacted swiftly and served as much needed lifelines for businesses to survive. The Ontario government also provided a range of support such as provincial tax rebates, loans and debt relief interventions, utility bill payment reductions, business grants, and low-interest loans, etc[@cite_reopening].

&nbsp; &nbsp; To help recover from the current recession, we at Petit Poll have partnered with the Ontario government to examine the effects and scope of COVID shutdown on restaurants. We conducted an experiment by selecting two statistically similar cities in Ontario and investigated how reopening the dine-in services for restaurants in one city has prevailed over the state of continued shutdown in another city. We compared unemployment rates, restaurant profitability, and the expected financial and social impact in both circumstances. The results showed that general satisfaction, a sense of purpose, hope, and profitability increase among the businesses that were allowed to re-open. The findings can be generalized to other similar businesses and their associated families and business partners in that city as well. In addition, the governmental support programs for small businesses in both cities contributed substantially to the social wellbeing and recovery pathway of all businesses. However, the higher comparative rate of COVID cases in the city under treatment resulted in general anxiety among residents, triggered the social health advocates and activists to criticize the government reopening effort, and undermined the projected benefits. We then came across the trade-off between opening or closing the public parts of restaurants - or other similar businesses, for which we suggested utilizing SSM (Soft System Methodology) to collect, consider and accommodate all residents' perspectives and potential viewpoints in society.

@@

**This experiment considered all Ontario restaurants as its target population. The frame population was decided by randomly selecting 100 restaurants from Brantford and 100 from Peterborough. To ensure the external validity of the experiment, we used the following steps to find these restaurants: **

@@

## Main Survey Results
```{r, include=FALSE}
# read in survey data
simulated_dataset<-read_csv(here('outputs/survey/simulated_restaurant_data.csv'))
```

```{r question3, fig.align='center', fig.cap="Layoff Situation due to COVID 19", fig.width=6, fig.height=3, echo= FALSE, message=FALSE, warning=FALSE}
data.frame(table(simulated_dataset$q3_layoff,simulated_dataset$q1_city))%>%
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
    geom_bar(stat = "identity",position = position_dodge(), width = 0.7)+
   theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Number of Laid-offs")+
  ggtitle("Layoff Situation due to COVID 19")+
  guides(fill=guide_legend(title="City"))

```

The above figure (Figure \@ref(fig:question3)) shows that restaurants in Peterborough had lower rates of layoff compared to restaurants in Brantford, suggesting families in Peterborough to have experienced better financial situations and probably less stress.

```{r q4, fig.align='center', fig.cap="Total Sales Volume Change Oct to Dec vs July to Sept in 2020", fig.width=6, fig.height=3, echo= FALSE, message=FALSE, warning=FALSE}

q4_table <- data.frame(table(simulated_dataset$q4_total_sales,simulated_dataset$q1_city))
data.frame(table(simulated_dataset$q4_total_sales,simulated_dataset$q1_city))%>%
  mutate(Var1 = fct_relevel(Var1,"Prefer not to say", "Decreased more than 10%","Decreased 5% ~ 10%","Roughly unchanged","Increased 5% ~ 10%","Increased more than 10%")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Total Sale Change")+
  ggtitle("Total Sales Volume Change Oct to Dec vs July to Sept in 2020")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()
```
From Figure \@ref(fig:q4), we notice a considerable increase in sales in Peterborough during the reopen period. This indicated that if restaurants were open, people would had used their dine-in services, regardless of potential pandemic risks. Besides, this result also corresponds with lower layoffs in Peterborough since restaurants need staff for their dine-in areas to operate.

In contrast, continuing shutdown restrictions during October to December on Brantford resulted in a slight decrease in sales volume for those restaurants. We propose possible reasons for the decrease in restaurant sales: (1) Residents in Brantford changed their mealing behaviors to cook more by themselves, instead of relying on meals from restaurants. (2) The continuing shutdown resulted in closures of other businesses, so these business staffs were not contributing for the restaurant sales anymore.

The increased costs of restaurants in Peterborough, as shown in Figure \@ref(fig:q5), demonstrated more activities in their restaurants. This change admits that the restaurants probably should have preserved more of their staff and consumed more utility or housekeeping materials to maintain their dine-in area. Although increased operating cost would usually impact the business unit negatively, from a broader view, these costs generate economic flow into local society and are vital for other business sectors to survive. So we consider the increased cost in restaurants during the reopening period as a good signal of economic wellbeing.

```{r q5, fig.align='center', fig.cap="Total Cost Volume Change Oct to Dec vs July to Sept in 2020", fig.width=6, fig.height=3, echo= FALSE, message=FALSE, warning=FALSE}

q5_table <- data.frame(table(simulated_dataset$q5_total_cost,simulated_dataset$q1_city))
q5_table <-  q5_table %>% add_row(Var1 = "Decreased by more than 25%", Var2 = "Peterborough", Freq = 0) %>%
  add_row(Var1 = "Decreased by more than 25%", Var2 = "Brantford", Freq = 0)
q5_table%>%
  mutate(Var1 = fct_relevel(Var1,"Prefer not to say","Decreased by more than 25%", "Decreased by 5% ~ 25%","Roughly unchanged","Increased by  5% ~ 25%","Increased by more than 25%")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Total Cost Change")+
  ggtitle("Total Cost Volume Change Oct to Dec vs July to Sept in 2020")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()

```
The following figure (\@ref(fig:q8)) demonstrates restaurant owners' opinions towards the expected survivability of their businesses. It is very unfortunate that restaurants in Brantford are not as positive to continue their businesses for longer times, indicating a stressful operating situation during these continuing shutdown period. While on the other hand, restaurants in Peterborough generally expect to survive longer, showing stronger confidence to continue their businesses after reopening.
```{r q8, fig.align='center', fig.cap="Expected Survival Time if Current Situation Continues", fig.width=6, fig.height=3, echo= FALSE, message=FALSE, warning=FALSE}
# Use

data.frame(table(simulated_dataset$q8_survive,simulated_dataset$q1_city))%>%
    mutate(Var1 = fct_relevel(Var1 ,"Unknown","Longer than 6 months","3 to 6 months","Less than 3 months")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Expected Survival Time")+
  ggtitle("Expected Survival Time if Current Situation Continues")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()
```

```{r q10, fig.align='center', fig.cap="Expected Recovery Time Back to pre-COVID Situation", fig.width=8, fig.height=4, echo= FALSE, message=FALSE, warning=FALSE}
# use
data.frame(table(simulated_dataset$q10_recover,simulated_dataset$q1_city,simulated_dataset$q2_type))%>%
  mutate(Var1 = fct_relevel(Var1 ,"Prefer not to say","More than 18 months","12 to 18 months","7 to 12 months","Six months or less")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge())+
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~Var3)+
  ylab("Number of responses")+
  xlab("Number of months")+
  ggtitle("Expected Recovery Time Back to pre-COVID Situation")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()

```
Although from Figure \@ref(fig:q8), we a pervasive gloom among the Brantford restaurants, Figure \@ref(fig:q10) illustrates that even if the pandemic ends immediately, many restaurants in Brantford, especially quick-service restaurants, have indicated that it is difficult to return to the state of operation before the COVID-19 outbreak. While for Peterborough, although some of them suspect the recovery will take more than a year to recoup with the previous situation, most of them indicated this period would be less than a year.

In addition, survey results about expected recovery time back to pre-COVID situation (Figure \@ref(fig:q10)) illustrate that even if the pandemic ends immediately, most restaurants in Brantford, especially quick-service restaurants, hold negative views to recover back to pre-COVID operating situation in a short period of time. For restaurants in Peterborough, despite some suspicions to recoup with the pre-COVID situation in more than a year, most of them expected to recovery in less than a year. Therefore, we believe that the continuing shutdown in Brantford have harmed expectations of future operation possibilities in Brantford, while the reopening in Peterborough has built confidence to resolve current operation difficulties.

@@

* Since the distance between Brantford and Peterborough is more than 200 kilometers apart, residents in Brantford are unlikely to travel such long distance solely for a dine-in meal. So the treatment of reopening Peterborough would not affect the mealing behavior in Brantford. However, residents in shutdown regions near Peterborough might go for dine-in meals, potentially increasing sales in Peterborough restaurants. The actual shutdown effect on restaurants could be overestimated due to possible increase in sales from nearby shutdown regions' residents.

* The mindful comparison between cities' similarities on employment, population density, etc. ensures the internal validity of our experiment so that we could accurately estimate the true impact of shutdown restrictions on restaurant. However, our shutdown effect evaluation may not be generalized to the entire population of interest, i.e. all restaurants in Ontario such as township regions or cosmopolitan city Toronto due to different levels of competitions, operating costs, etc.

## Concerns about Sampling Method

@@

## Survey Results and Policy Implication

According to survey questions 3-17 in section 2, Peterborough restaurants indicated to be more cheerful to carry on with their businesses in future. They found the COVID-19 situation remediable, demonstrated the desire to further invest in their career, and revealed to be an overall happier society compared to Brantford restaurant. As Dr. Ed Diener, a professor of psychology at the University of Virginia and the University of Utah, and a senior scientist for the Gallup Organization stated in his interview with Sherif Arafa(https://greatergood.berkeley.edu/article/item/why_governments_should_care_more_about_happiness), “The deep and long-lasting sources of happiness contribute to what we call sustainable happiness, which is lasting, not just momentary”. He continues “we find that sustainable happiness leads to:
·       Better health and longevity: Happy people live longer and experience better health.
·       Superior work performance, especially organizational citizenship.
·       More supportive social relationships—for example, being less likely to get divorced.
·       Better citizenship and more “prosocial” behaviors—for example, being more likely to help others, volunteer, or donate money to charity.
·       Better mental health and resilience when confronted by stressful events—for example, being less likely to suffer from mental health problems such as depression, and more likely to bounce back after something bad happens.
The benefits of sustainable happiness are broad and important. They cannot be ignored. To repeat: Happy people do not just feel better, but they function better in achieving the things we value. They help society. They help those around them to also lead high-quality lives. Happy people are not uncaring, selfish people. They have the energy and desire to work on societal problems as much as or more than others”. According to our finding in the above survey results, we believe reopening of the restaurants in Peterborough has resulted in “sustainable happiness” among restaurant owners in this city. Further to that, we didn’t find a substantial reason to not generalize these results to other similar businesses who need to provide physical areas to serve their clients respectively. Adhering to strict preventing regulations, this would include professional services like clinical treatments, open area sports and other applicable safe businesses. We also argue that the  resulted happiness, which was proactively forethought and managed by the Government of Canada by offering supportive programs such as Canada Emergency Wage Subsidy (CEWS), Canada Emergency Loan Subsidy (CELS), business grants and low interest loans, will not only decrease the rates of depression, suicide, frustration, crime, and other unprecedented social issues among this society, it will definitely leave a sense of trust, confidence and faith among the adolescence who usually start their career with a pat-time student work in restaurants. According to (https://www.sciencedaily.com/releases/2011/08/110822091859.htm), “the relative intensity of these emotions (happiness and depression among adolescents) is also important. The odds of drug use were notably lower for youth who reported that they were more often happy than depressed and were substantially higher for those who indicated that they were more depressed than happy”. To this point, we strongly believe that the value of our understudy treatment “the reopening restaurants in Peterborough” has been realized and will continue in future.

On the other hand, the result of covid cases (((((reference???))))))) revealed a new safety related concern. The higher rate of Corona incidents, fear and mortality among Peterborough people, compared to the residents of Brantford, take us to the emergence of a new argument, “how do you find the overall happiness among people in a city, not solely among the restaurant owners?”. Undoubtedly, more CORONA related cases fold up happiness and spread depression among society, which literally criticises the beneficial effects of above mentioned governmental supportive programs.

Do we believe in absolute “safety first”? If we do, no doubt about keeping restaurants, and probably all other similar business activities, closed until the pandemic ends. And, the government should invest in other programs to insert hope, administer happiness and hinder the spread of depression. However, a controversial discussion would be “should considerate people pay for the negligence of defiant citizens who parade time to time at Yonge and Bloor streets in Toronto (in front of my apartment) to criticize wearing masks?”. Not good enough, some suspect the Corona vaccine contains microchips or 3D-camera to spy and scan their body for intelligent services!

The goal of this paper is to report the dynamics of an intervention and to analyze how the results of a treatment - reopening the restaurants - would happen if in place. We observe a trade-off between the prosperity and hopefulness among the businesses in one city versus the lower number of COVID cases throughout the other city. We believe this would be a controversial discussion between citizens, activists and the government. At this situation, we recommend the “Soft System Methodology” (SSM) which “uses “systems thinking” in a cycle of action research, learning and reflection to help understand the various perceptions that exist in the minds of the different people involved in (one) situation” (http://www.construction-innovation.info/images/pdfs/Research_library/ResearchLibraryA/Refereed_Conference_papers/Five_Case_Studies.pdf). SSM contributes to better knowledge management, leading to better understanding of different existing viewpoints and also insights to different perspectives of communities. Gathering more information and context knowledge would more or less result in taking the most desired and applicable approach to enforce decisions and to better respond to the citizen’s needs.

### 2021-02-24 scripts/2178_chart_2.Rmd
Commit for discussion

#Q3 layoff
```{r}
data.frame(table(simulated_dataset$q3_layoff,simulated_dataset$q1_city))%>%
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
    geom_bar(stat = "identity",position = position_dodge(), width = 0.7)+
   theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Number of Laid-offs")+
  ggtitle("Layoff Situation due to COVID 19")+

  guides(fill=guide_legend(title="City"))

@@

pie(p_table$Freq,labels = label_p,col=c("#7FC97F","#BEAED4", "#FDC086", "#FFFF99", "#386CB0" ),
    main = "Peterborough: Total sales volume during Oct to Dec 2020 compared to July to Sept 2020",cex=0.6,cex.main=1)

#Brantford
b_table <- data.frame(table(simulated_dataset[simulated_dataset$q1_city=="Brantford",]$q4_total_sales))%>%
  mutate(pct=round(Freq/sum(Freq)*100,2))

@@

#Brantford
b_table_cost<- data.frame(table(simulated_dataset[simulated_dataset$q1_city=="Brantford",]$q5_total_cost))%>%
  mutate(pct=round(Freq/sum(Freq)*100,2))

@@

```{r}
# use
data.frame(table(simulated_dataset$q14_loans,simulated_dataset$q1_city))%>%
  mutate(Var1 = fct_relevel(Var1 ,"Prefer not to say","More than $50,000","$40,000 ~ $50,000","$20,000 ~ $40,000","$0 ~ $20,000")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
    geom_bar(stat = "identity",position = position_dodge())+
  ylab("Number of responses")+
  xlab("Grant value")+
  ggtitle("How much loans have you received for the COVID-19 related reasons?")+
  guides(fill=guide_legend(title="City"))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 0))+
  coord_flip()

@@

# Q15

>>>>

#Q3 layoff
```{r}
data.frame(table(simulated_dataset$q3_layoff,simulated_dataset$q1_city))%>%
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
    geom_bar(stat = "identity",position = position_dodge(), width = 0.7)+
   theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Number of Laid-offs")+
  ggtitle("Layoff Situation due to COVID 19")+
  guides(fill=guide_legend(title="City"))

@@

pie(p_table$Freq,labels = label_p,col=c("#7FC97F","#BEAED4", "#FDC086", "#FFFF99", "#386CB0" ),
    main = "Peterborough: Total sales volume during Oct to Dec 2020 compared to July to Sept 2020",cex=0.6,cex.main=1)
```

```{r}
# use
q4_table <- data.frame(table(simulated_dataset$q4_total_sales,simulated_dataset$q1_city))
data.frame(table(simulated_dataset$q4_total_sales,simulated_dataset$q1_city))%>%
  mutate(Var1 = fct_relevel(Var1,"Prefer not to say", "Decreased more than 10%","Decreased 5% ~ 10%","Roughly unchanged","Increased 5% ~ 10%","Increased more than 10%")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Total Sale Change")+
  ggtitle("Total Sales Volume Change Oct to Dec vs July to Sept in 2020")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()
```

```{r}
#Brantford
b_table <- data.frame(table(simulated_dataset[simulated_dataset$q1_city=="Brantford",]$q4_total_sales))%>%
  mutate(pct=round(Freq/sum(Freq)*100,2))

@@

```{r}
# use
q5_table <- data.frame(table(simulated_dataset$q5_total_cost,simulated_dataset$q1_city))
q5_table <-  q5_table %>% add_row(Var1 = "Decreased by more than 25%", Var2 = "Peterborough", Freq = 0) %>%
  add_row(Var1 = "Decreased by more than 25%", Var2 = "Brantford", Freq = 0)
q5_table%>%
  mutate(Var1 = fct_relevel(Var1,"Prefer not to say","Decreased by more than 25%", "Decreased by 5% ~ 25%","Roughly unchanged","Increased by  5% ~ 25%","Increased by more than 25%")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
  theme_minimal()+
  geom_bar(stat = "identity",width = 0.8, position = position_dodge()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Number of Responses")+
  xlab("Total Cost Change")+
  ggtitle("Total Cost Volume Change Oct to Dec vs July to Sept in 2020")+
  guides(fill=guide_legend(title="City"))+
  coord_flip()
```

```{r}
#Brantford
b_table_cost<- data.frame(table(simulated_dataset[simulated_dataset$q1_city=="Brantford",]$q5_total_cost))%>%
  mutate(pct=round(Freq/sum(Freq)*100,2))

@@

```{r}
# use
data.frame(table(simulated_dataset$q14_loans,simulated_dataset$q1_city))%>%
  mutate(Var1 = fct_relevel(Var1 ,"Prefer not to say","More than $50,000","$40,000 ~ $50,000","$20,000 ~ $40,000","$0 ~ $20,000")) %>% 
  ggplot(aes(Var1, Freq,fill=Var2)) +
    geom_bar(stat = "identity", width = 0.8, position = position_dodge())+
  ylab("Number of responses")+
  xlab("Grant value")+
  ggtitle("How much loans have you received for the COVID-19 related reasons?")+
  guides(fill=guide_legend(title="City"))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 0))+
  coord_flip()

@@

# Q11
